---
title: "Simple Drake Example"
author: "Simon Knapp"
date: "17 March 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Interesting Code

```{r plan, collapse=TRUE}
#----------------------------------------------------------
# intersting functions
#----------------------------------------------------------
average.treatment.effect <- function(y, t, p) {
  t <- as.numeric(as.character(t))
  den <- sum(t/p + (1-t)/(1-p))
  sum(t*y/p - (1-t)*y/(1-p)) / den
}

#----------------------------------------------------------
# drake plan
#----------------------------------------------------------
library(drake)

plan <- drake_plan(
  all.data = read.csv(file_in('test.csv')),
  
  ps.model = glm(
    treatment ~ age,
    data=all.data,
    family=binomial),
  
  ps = predict(ps.model, type='response'),
  t.inds = with(all.data, as.numeric(as.character(treatment))),
  ips = with(all.data, t.inds/ps + (1-t.inds)/(1-ps)),
  
  all.in.mod.with.ps = lm(
    y ~ treatment,
    data=all.data,
    weights=ips),
  
  all.in.mod.with.no.ps = lm(
    y ~ treatment,
    data=all.data),
  
  all.in.mod.with.age = lm(
    y ~ treatment + age,
    data=all.data),
  
  ce = coef(all.in.mod.with.ps)['treatment'],
  ce.biased = coef(all.in.mod.with.no.ps)['treatment'],
  
  ce.age = coef(all.in.mod.with.age)['treatment'],
  
  ate = average.treatment.effect(all.data$y, all.data$treatment, ps),
  ate.biased = average.treatment.effect(all.data$y, all.data$treatment, 0.5)
)

#----------------------------------------------------------
# make the plan
#----------------------------------------------------------
make(plan)
```

```{r utils, echo = FALSE}
#----------------------------------------------------------
# utilities
#----------------------------------------------------------
fmt <- function(x) sprintf(x, fmt='%0.2f')
```

## The plan as a DAG

```{r dag}
config <- drake_config(plan)
vis_drake_graph(config)
```


## Some Estimators (using the results)

Note that the true ATE/causal effect is zero.


### Design Based Causal Effects

- Including Ps, the (ubiased) ATE is `r fmt(readd(ate))`

- Ignoring PS, the (biased) ATE is `r fmt(readd(ate.biased))`.
  
### Model Based Causal Effects

- The (unbiased) causal effect using a weighted least squares estimate in the
  incorrect regression model is `r fmt(readd(ce))`.
  
- The (biased) causal effect using a non-weighted least squares
  estimate in the incorrect regression model is `r fmt(readd(ce.biased))`.
  
- The (unbiased) causal effect using a non-weighted least squares and the
  correct regression model is `r fmt(readd(ce.age))`.